Differentiable Optics with {\Large $\partial$}Lux: I - Deep Calibration of Flat Field and Phase Retrieval with Automatic Differentiation

### 2022-12-16 src/tex/ms.tex
[showyourwork] overleaf sync

This operation is then performed over 15 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths 545nm, 595nm and 645nm.. This total model requires the calculation of 30 individual \ac{psf}s over a $1024\times1024$ detector, which takes a total of \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 CPU. The data set is generated by taking this set of images and applying Poisson photon noise to the \ac{psf}s and additive Gaussian noise to every pixel as detector read-noise with mean 10 and deviation 2.5 Using this image we recover stellar positions and fluxes, optical aberrations, and individual pixel sensitivities. This gives us 30 stellar positional parameters (15 RA and DEC each), 15 stellar flux parameters, 7 optical aberration parameters, and 1,048,576 detector parameters: the dimensionality of the problem is vastly dominated by the \ac{prf} parameters, which is why this problem is made tractable only by using autodiff.

@@

The optimisation is performed for a total of 100 epochs, with each gradient calculation taking \textcolor{red}{showyourwork variable} seconds to evaluate on an Apple M1 Max CPU, leading to a total optimisation time of \textcolor{red}{showyourwork variable} minutes.

This optimisation strategy performs well, with all of the parameters being recovered down to the noise level. This can be seen in Figure~\ref{fig:data_residual} with the right panel showing a sample residual of the data and model after optimisation, which is clearly only noise. Correlations and residual plots of the positions and fluxes after the optimisation are shown in Figure~\ref{fig:astro_params}. The 1-$sigma$ error bars indicate a poor recovery, however these are only marginalised over positions, fluxes and aberration, not the \ac{prf} which is the dominant noise source. These parameters are recovered to the expected distribution when performed without modelling the \ac{prf}. Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of 
% \textcolor{red}{showyourwork variable}
\variable{output/rms_opd_resid.txt} nm RMS. 
Finally, Figure~\ref{fig:flat_field} shows a correlation plot of the true and recovered sensitivity values and the corresponding residual histogram. The points of the correlation plot are colour-coded with the total counts for each pixel, showing as expected that pixels with higher photon counts having a better recovery of the PRF since these pixels have a greater \ac{snr}. Overall the parameters are very well recovered down to the noise level.

>>>>

This operation is then performed over 15 stars with uniform 100\,nm bandpass spectra sampled at three wavelengths 545nm, 595nm and 645nm.. This total model requires the calculation of 30 individual \ac{psf}s over a $1024\times1024$ detector, which takes a total of $\sim$ 1.25 seconds to evaluate on an Apple M1 CPU. The data set is generated by taking this set of images and applying Poisson photon noise to the \ac{psf}s and additive Gaussian noise to every pixel as detector read-noise with mean 10 and deviation 2.5 Using this image we recover stellar positions and fluxes, optical aberrations, and individual pixel sensitivities. This gives us 30 stellar positional parameters (15 RA and DEC each), 15 stellar flux parameters, 7 optical aberration parameters, and 1,048,576 detector parameters: the dimensionality of the problem is vastly dominated by the \ac{prf} parameters, which is why this problem is made tractable only by using autodiff.

@@

The optimisation is performed for a total of 100 epochs, with each gradient calculation taking $\sim$ 2.75 seconds to evaluate on an Apple M1 Max CPU, leading to a total optimisation time of $\sim$ 4.5 minutes.

This optimisation strategy performs well, with all of the parameters being recovered down to the noise level. This can be seen in Figure~\ref{fig:data_residual} with the right panel showing a sample residual of the data and model after optimisation, which is clearly only noise. Correlations and residual plots of the positions and fluxes after the optimisation are shown in Figure~\ref{fig:astro_params}. The 1-$sigma$ error bars 
are generated by harnessing the differentiablility of the model by calculating the covaraince matrix under the Laplace approximation, defined as the negative matrix inverse of the hessian of the likelihood function evaluated at the maximum likelihood estimate, ie the final recovered values. These indicate a poor recovery, however they are only marginalised over positions, fluxes and aberration, not the \ac{prf} which is the dominant noise source as its high dimensionality makes marginalisation computationally intractable. These parameters are recovered to the expected distribution when performed without modelling the \ac{prf}. Figure~\ref{fig:aberrations} shows the recovery of the optical aberrations, with a final residual aberration of 
% \textcolor{red}{showyourwork variable}
\variable{output/rms_opd_resid.txt} nm RMS. 
Finally, Figure~\ref{fig:flat_field} shows a correlation plot of the true and recovered sensitivity values and the corresponding residual histogram. The points of the correlation plot are colour-coded with the total counts for each pixel, showing as expected that pixels with higher photon counts having a better recovery of the PRF since these pixels have a greater \ac{snr}. Overall the parameters are very well recovered down to the noise level.
